Hypertension and microbiome

### 2019-01-23 articletwo.Rmd
Minor

Importing libraries:

@@

```{r Initialize docx, include = FALSE, cache = FALSE}
doc <- read_docx(path = "style/articlestyle.docx") %>%
  body_remove()

dir.create("rds", showWarnings = FALSE)
dir.create("cache", showWarnings = FALSE)

text.bold <- fp_text(font.size = 12, bold = TRUE, font.family = "Arial")
text.normal <- fp_text(font.size = 12, font.family = "Arial")
```

@@

```{r Write characteristics}
tbl1head <- "Characteristics of the study sample."
tbl1foot <- "Continuous variables are presented as mean (standard deviation). Categorical variables reported as absolute and relative frequencies. BP indicates blood pressure, BMI body mass index, RAS renin-angiotensin system."
writetable(doc, tbl1, number = 1, tbl1head, tbl1foot)
print(doc, target = "articletwo.docx")
```

@@

adonis.species  <- readRDS("rds/adonis.species.rds")
beta.diversity <- adonis.species %>%
    data.table::rbindlist(id = "model_name") %>%
    dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION")) %>%
    dplyr::select(response, R2)

@@

@@

```{r Write docx to file, cache = FALSE, results = FALSE, include = FALSE}
print(doc, target = "articletwo.docx")
```

>>>>

# Running at

Calculations ran at

```{r Running at}
if (!exists('now')) {
    now <- format(Sys.time(), '%Y%m%d-%H%M%S')
}
now
```

# Importing libraries:

@@

```{r Initialize docx, include = FALSE, cache = FALSE}
doc <- read_docx(path = "style/articlestyle.docx") %>%
  body_remove()
dir.create("rds", showWarnings = FALSE)
dir.create("cache", showWarnings = FALSE)
```

@@

```{r Write characteristics}
tbl1head <- "Characteristics of the study sample."
tbl1foot <- "Continuous variables are presented as mean (standard deviation). Categorical variables reported as absolute and relative frequencies. BP indicates blood pressure, BMI body mass index, RAS renin-angiotensin system."
writetable(doc, tbl1, number = 1, tbl1head, tbl1foot)
```

## Alpha & Beta plot (Figure 1.)

@@

adonis.species  <- readRDS("rds/adonis.species.rds")

beta.diversity <- adonis.species %>%
    data.table::rbindlist(id = "model_name") %>%
    dplyr::filter(term %in% c("MAP", "SYSTM", "DIASM", "PULSEPRESSURE", "HYPERTENSION"))

                                        #%>%
    dplyr::select(response, R2)

@@

## Principal component analysis (Figure 2.)

@@

```{r Write docx to file, cache = FALSE, results = FALSE, include = FALSE}
print(doc, target = paste0("report/articletwo-", now, ".docx"))
```

### 2019-01-23 rrbiome.Rmd
Minor

# Function definitions

```{r mini functions}
replace.brackets <- function (genus) gsub('(.*) \\(+(.+)\\)', '\\1.\\2', genus)
```

```{r functions}
filter.phenotype.data <- function(pseq,
                                  included = c("DNA_OK", "MEN", "BL_AGE", "SYSTM", "DIASM", "BP_TREAT", "BMI",
                                               "CURR_SMOKE", "DIAB", "BL_USE_RX_C09", "Q57X", "BL_USE_RX_C03",
                                               "BL_USE_RX_C07", "BL_USE_RX_C08"))
    meta(pseq) %>%
        tibble::rownames_to_column(var = "rowname") %>%
        dplyr::select(rowname, included) %>% 
        dplyr::mutate(ANYDRUG = factor(ifelse(BL_USE_RX_C03  == 1 | BL_USE_RX_C07 == 1 |
                                       BL_USE_RX_C08 == 1  | BL_USE_RX_C09 == 1, 1, 0)),
                      ANYEXERCICE = factor(ifelse(Q57X == 1, 0, 1)),
                      PULSEPRESSURE = SYSTM - DIASM,
                      HYPERTENSION = factor(ifelse(SYSTM >= 140 | DIASM >= 90 | ANYDRUG == 1, 1, 0)),
                      SEX = factor(ifelse(MEN == "Female", 1, 0)),
                      MAP = 2./3.*DIASM + 1./3.*SYSTM) %>%
         stats::na.omit() %>%
        dplyr::select(-MEN, -DNA_OK) %>%
        tibble::remove_rownames() %>%
        tibble::column_to_rownames(var = "rowname")

meta.merge.alphadiversity <- function(pseq, index = "shannon") {
    alphadiversity  <- microbiome::global(pseq, index = index)
    base::merge(meta(pseq), alphadiversity, by=0, all=TRUE) %>%
        dplyr::rename(Sample_ID = Row.names) %>%
            dplyr::mutate(BL_AGE_GROUP = group_age(BL_AGE))
}

calculate.beta.matrix <- function(pseq) {
    compositional.profile <- microbiome::transform(pseq, 'compositional')
    otu <- microbiome::abundances(compositional.profile)
    bray.dist.m <- vegan::vegdist(t(otu), method="bray")
    dist.matrix <- as.matrix(bray.dist.m)
    attr(dist.matrix, "method") <- "bray"
    dist.matrix
}
calculateglm <- function(dset,
                         responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
                                          model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
                         min_n_for_continuous = 10,
                         modelstring = "%s ~ BL_AGE + SEX") {
    glmlist <- lapply(responses, function(response) {
        fo.family <- ifelse(length(unique(pull(dset, response))) > min_n_for_continuous, stats::gaussian, stats::binomial)
        fo <- sprintf(modelstring, response)
        stats::glm(formula = as.formula(fo), family = fo.family, data = dset) %>%
            broom::tidy() %>%
                dplyr::filter(grepl("diversities_", term)) %>%
                dplyr::mutate(response = response, fo = fo) %>%
                dplyr::select(-term)
    })
    data.table::rbindlist(glmlist, id = "model_name") %>%
        as.data.frame %>%
        dplyr::mutate(conf.low = estimate - qnorm(1- 0.05/2) * std.error,
                      conf.high = estimate + qnorm(1- 0.05/2) * std.error)
}

calculateadonis <- function(dset,
                            matrix,
                            responses = list(model_1 = "MAP", model_2 = "SYSTM", model_3 = "DIASM",
                                          model_4 = "PULSEPRESSURE", model_5 = "HYPERTENSION"),
                            covariates = c("BL_AGE", "SEX"),
                            npermutations = 99,
                            maxcores = 100) {
    mclapply(responses, function(response) {
	fo <- sprintf("matrix ~ %s + %s", response, paste(covariates, collapse = " + "))
        ad <- adonis(as.formula(fo), data = dset, permutations = npermutations)
        ad$aov.tab %>%
            as.data.frame %>%
            tibble::rownames_to_column(var = "term") %>%
            dplyr::mutate(response = response)
      }, mc.cores = min(maxcores, length(responses)))
}

prepare.maaslin <- function (pseq,
                             tsvfile = "merged_phfinrisk_genus.tsv",
                             conffile = "maaslin_config.config",
                             includedvars = c("BL_AGE", "BMI", "SEX"),
                             readpclrows = "Archaea.Archaea-",
                             core = FALSE,
                             core.detection = 0.001,
                             core.prevalence = 0.01) {
    pseq.comp <- microbiome::transform(pseq, "compositional")
    if (core) {
         pseq.comp  <- microbiome::core(pseq.comp, detection = core.detection, prevalence = core.prevalence)
    }
    pheno <- t(microbiome::meta(phyloseq::sample_data(pseq.comp)))
    sample.ids <- colnames(pheno)
    abud.rel <- abundances(pseq.comp)
    rownames(abud.rel) <- sapply(rownames(abud.rel), replace.brackets)
    write.table(t(rbind(sample.ids, pheno, abud.rel)), file = tsvfile, sep='\t', quote=FALSE, row.names=FALSE, col.names=TRUE)
    data <- file(conffile , open='wt')
    pcl.contents <- c("Matrix: Metadata",
                      paste(c("Read_PCL_Rows: ", paste(includedvars, collapse=",")), collapse = ""),
                      "",
                      "Matrix: Adundance",
                      paste("Read_PCL_Rows:", readpclrows))
    writeLines(pcl.contents, con=data)
    close(data)
}

 merge.pheno.abu <- function(pseq,
                             core = FALSE,
                             core.detection = 0.001,
                             core.prevalence = 0.01) {
     pseq.comp <- microbiome::transform(pseq, "compositional")
     if (core) {
         pseq.comp  <- microbiome::core(pseq.comp, detection = core.detection, prevalence = core.prevalence)
    }
     pheno <- microbiome::meta(phyloseq::sample_data(pseq.comp))
     sample.ids <- rownames(pheno)
     abud.rel <- abundances(pseq.comp)
     rownames(abud.rel) <- sapply(rownames(abud.rel), replace.brackets)
     cbind(pheno, t(abud.rel))
}

```

>>>>

## RR biome functions

```{r RR biome functions, code=readLines("articletwo-rrbiome.R")}
source("articletwo-rrbiome.R")
```
